Methods for Analyzing Models of Rod-Shaped Bacteria

### 2025-07-10 paper/main.tex
adjust formulation
add idea for discussion

Bacteria exist in many physical shapes such as spheroidal, rod-shaped or spiral
\cite{Zapun2008,Young2006}.
The last decades have fundamentally challenged and changed our understanding of how bacteria retain
their shape.
Rod-shaped bacteria can grow by extending their circular part, or by inserting new material at the
tip of the rod.
Species such as E.coli and B.subtilis~\cite{Errington2020} fall under the first category while
S.pombe represents the latter.
The material is inserted in small bursts on the nanoscale in forms of patches, bands or
hoops~\cite{DePedro2003}.
Despite their differences in wall-size, B.subtilis and E.coli follow growth rules which are
comparable with respect to the extension of the rod~\cite{Chang2014}.

@@

\begin{itemize}
    \item Model does not capture every aspect of rod-shaped bacteria. Needs to extended (see our
        review).
    \begin{enumerate}
        \item polar interactions
        \item differentiation for van Gogh bundles; surfactin-producing, matrix-producing cells.
        \item Proper friction; this might have a huge implication on the estimation of
            parameters~\cite{Grant2014}.
            This paper has already implemented this~\cite{Doumic2020}.
            We can build on top of that.
        \item Extracellular Reactions~\cite{Li2025}
        \item see supplement for full list \ref{table:simulation-aspects-supplement}
    \end{enumerate}
    \item Image generation; Produce high-quality near-realistic microscopic images with cell-masks
        from scratch with mechanistic model; train cell-segmentation and more importantly
        cell-tracking algorithms with that.
\end{itemize}

>>>>

Bacteria exist in many physical shapes such as spheroidal, rod-shaped or spiral
\cite{Zapun2008,Young2006}.
The last decades have fundamentally challenged and changed our understanding of how bacteria retain
their shape.
Rod-shaped bacteria can grow by extending their circular part, or by inserting new material at the
tip of the rod.
Species such as E.coli and B.subtilis~\cite{Errington2020} fall under the first category while
S.pombe represents the latter.
New material is inserted in small bursts on the nanoscale in forms of patches, bands or
hoops~\cite{DePedro2003}.
Despite their differences in wall-size, B.subtilis and E.coli follow growth rules which are
comparable with respect to the extension of the rod~\cite{Chang2014}.

@@

\begin{itemize}
    \item Model does not capture every aspect of rod-shaped bacteria. Needs to extended (see our
        review).
    \begin{enumerate}
        \item polar interactions
        \item differentiation for van Gogh bundles; surfactin-producing, matrix-producing cells.
        \item Proper friction; this might have a huge implication on the estimation of
            parameters~\cite{Grant2014}.
            This paper has already implemented this~\cite{Doumic2020}.
            We can build on top of that.
        \item Extracellular Reactions~\cite{Li2025}
        \item see supplement for full list \ref{table:simulation-aspects-supplement}
    \end{enumerate}
    \item Data Extraction: Skeletonization Algorithm ultimately not good enough for some scenarios
        (i.e. overlapping of cells, endpoints can sometimes be "too short")
    \begin{itemize}
        \item Possibly use AI for new algorithm
    \end{itemize}
    \item Image generation; Produce high-quality near-realistic microscopic images with cell-masks
        from scratch with mechanistic model; train cell-segmentation and more importantly
        cell-tracking algorithms with that.
\end{itemize}
